---
title: "Pence Kaine EDA"
output: pdf_document
---

```{r load-packages, message = FALSE, echo=FALSE}
library(rvest)
library(dplyr)
library(tidyr)
library(knitr)
```

```{r load-data}
checkdata = read.csv("data/checkdata.csv")
pencekaine = read.csv("data/pence_kaine_2016.csv")
```

### ClaimBuster Ratings

```{r arrange-checked, echo=FALSE}
checked_claims = checkdata %>% 
  filter(Checked. == "1") %>% 
  arrange(desc(Ratings))

checked_claims %>% 
  group_by(Claimant) %>% 
  summarize(avgrating = mean(Ratings)) %>%
  kable()
```

Lowest ClaimBuster score of a claim that was checked: 0.12

The highest ClaimBuster score of a claim that was checked: 0.71

Total claims checked: 112

Average ClaimbBuster score Mike Pence: *\~.42*

Average ClaimBuster score Tim Kaine: *\~.38*

**On average Tim Kaine's claims are slightly less checkable (0.04 / 4%).**

### Claims Checked and Unchecked

```{r claimant-info, echo=FALSE}
checkdata %>% 
  group_by(Claimant) %>% 
  count() %>%
  kable()

checked_claims %>%
  group_by(Claimant) %>% 
  count() %>%
  kable()
```

Mike Pence had a total of 542 factual claims. 53 of his factual claims were checked *\~9.8%*

Tim Kaine had a total of 582 factual claims. 59 of his factual claims were checked *\~10%*

**No major bias by fact checkers according to candidate** Note: may be worth running a t-test on this to confirm no statistical significance.

### Repeated Fact-Checks

```{r times-checked, eval=FALSE, echo=FALSE}
checked_claims %>% 
  group_by(times.fact.checked) %>% 
  count(Claimant) %>%
  kable()
```

```{r times-checked 2, echo=FALSE}
checked_claims %>%
  count(times.fact.checked, Claimant) %>%
  pivot_wider(id_cols = c(times.fact.checked, Claimant),
              names_from = times.fact.checked, 
              values_from = n,
              values_fill = 0) %>%
  kable()
```

As expected (knowing that each claimant was fact-checked a roughly equal amount of times) both claimants have similar amounts of claims checked repeatedly.

The only *times.fact.checked* category that has different numbers of claims checked for each Claimant is *2* where Mike Pence had *11* checked claims and Tim Kaine has *16* checked claims.

### Differences between Publishers

```{r fact-checkers, echo=FALSE, message=FALSE}
PF = checked_claims %>%
  filter(PF == "1") %>%
  group_by(Claimant) %>%
  summarize(PFcheck = n())

WP = checked_claims %>%
  filter(WP == "1") %>%
  group_by(Claimant) %>%
  summarize(WPcheck = n())

FC = checked_claims %>%
  filter(FC == "1") %>%
  group_by(Claimant) %>%
  summarize(FCcheck = n())

NYT = checked_claims %>%
  filter(NYT == "1") %>%
  group_by(Claimant) %>%
  summarize(NYTcheck = n())

CNN = checked_claims %>%
  filter(CNN == "1") %>%
  group_by(Claimant) %>%
  summarize(CNNcheck = n())

AP = checked_claims %>%
  filter(AP == "1") %>%
  group_by(Claimant) %>%
  summarize(APcheck = n())

TG = checked_claims %>%
  filter(TG == "1") %>%
  group_by(Claimant) %>%
  summarize(TGcheck = n())

checkers = left_join(PF, left_join(WP
                                   , left_join(FC
                                               , left_join(NYT
                                                           , left_join(CNN
                                                                       , left_join(AP, TG, 
                                                                                   by = "Claimant"))))))
checkers %>% kable()                                                                       
```

1/7 checker checked both candidates equally (TG)

3/7 checkers checked Pence more than Kaine (NYT, CNN, AP)

3/7 checkers checked Kaine more than Pence (PF, WP, FC)

**No apparent partisan bias by news sources for one candidate over the other**

### With 0.2 ClaimBuster Threshold

```{r 0.2-checked}
`0.2checked` = checked_claims %>% 
  filter(Ratings >= 0.2)

`0.2data` = checkdata %>% 
  filter(Ratings >= 0.2)

`0.2checked` %>%
  group_by(Claimant) %>%
  summarize(avgrating = mean(Ratings))

`0.2data` %>%
  group_by(Claimant) %>%
  count()

`0.2checked` %>%
  group_by(Claimant) %>%
  count()
```

Average claimbuster score Mike Pence: *\~.44*

Average claimbuster score Tim Kaine: *\~.39*

**On average Tim Kaine's claims are slightly less checkable (0.05 / 5%).**

Mike Pence had a total of 358 factual claims. 50 of his factual claims were checked *\~14%*

Tim Kaine had a total of 402 factual claims. 57 of his factual claims were checked *\~14%*

**No major bias by fact checkers according to candidate** Note: may be worth running a t-test on this to confirm no statistical significance.

```{r fact-checkers.2}
PF.2 = `0.2checked` %>%
  filter(PF == "1") %>%
  group_by(Claimant) %>%
  summarize(PFcheck = n())

WP.2 = `0.2checked` %>%
  filter(WP == "1") %>%
  group_by(Claimant) %>%
  summarize(WPcheck = n())

FC.2 = `0.2checked` %>%
  filter(FC == "1") %>%
  group_by(Claimant) %>%
  summarize(FCcheck = n())

NYT.2 = `0.2checked` %>%
  filter(NYT == "1") %>%
  group_by(Claimant) %>%
  summarize(NYTcheck = n())

CNN.2 = `0.2checked` %>%
  filter(CNN == "1") %>%
  group_by(Claimant) %>%
  summarize(CNNcheck = n())

AP.2 = `0.2checked` %>%
  filter(AP == "1") %>%
  group_by(Claimant) %>%
  summarize(APcheck = n())

TG.2 = `0.2checked` %>%
  filter(TG == "1") %>%
  group_by(Claimant) %>%
  summarize(TGcheck = n())

checkers.2 = left_join(PF.2
                     , left_join(WP.2
                                 , left_join(FC.2
                                             , left_join(NYT.2
                                                       , left_join(CNN.2
                                                                  , left_join(AP.2, TG.2, 
                                                                             by = "Claimant"))))))
checkers.2                                                                       
```

3/7 checkers checked Pence more than Kaine (NYT, CNN, AP)

4/7 checkers checked Kaine more than Pence (PF, WP, FC, TG)

**No apparent partisan bias by news sources for one candidate over the other**
