---
title: "Pence Kaine EDA"
output: pdf_document
---

```{r load-packages, message = FALSE}
library(rvest)
library(dplyr)
library(tidyr)
```


```{r load-data}
checkdata = read.csv("data/checkdata.csv")
pencekaine = read.csv("data/pence_kaine_2016.csv")
```

```{r arrange-checked}
checked_claims = checkdata %>% 
  filter(Checked. == "1") %>% 
  arrange(Ratings)
```
Lowest claimbuster score of a claim that was checked: 0.12 
The highest claimbuster score of a claim that was checked: 0.71 
Total claims checked: 112

```{r claimant-info}
checkdata %>% 
  group_by(Claimant) %>% 
  count()

checked_claims %>%
  group_by(Claimant) %>% 
  count()
```
Mike Pence had a total of 542 factual claims. 53 of his factual claims were
checked *~9.8%* 
Tim Kaine had a total of 582 factual claims. 59 of his factual claims were 
checked *~10%* 
**No major bias by fact checkers according to candidate** Note: may be worth 
running a t-test on this to confirm no statistical significance. 

```{r times-checked}
checked_claims %>% 
  group_by(times.fact.checked) %>% 
  count(Claimant)
```
As expected (knowing that each claimant was fact-checked a roughly equal amount
of times) both claimants have similar amounts of claims checked repeatedly. 
The only *times.fact.checked* category that has different numbers of claims
checked for each Claimant is *2* where Mike Pence had *11* checked claims and 
Tim Kaine has *16* checked claims. 

```{r fact-checkers}
PF = checked_claims %>%
  filter(PF == "1") %>%
  group_by(Claimant) %>%
  summarize(PFcheck = n())

WP = checked_claims %>%
  filter(WP == "1") %>%
  group_by(Claimant) %>%
  summarize(WPcheck = n())

FC = checked_claims %>%
  filter(FC == "1") %>%
  group_by(Claimant) %>%
  summarize(FCcheck = n())

NYT = checked_claims %>%
  filter(NYT == "1") %>%
  group_by(Claimant) %>%
  summarize(NYTcheck = n())

CNN = checked_claims %>%
  filter(CNN == "1") %>%
  group_by(Claimant) %>%
  summarize(CNNcheck = n())

AP = checked_claims %>%
  filter(AP == "1") %>%
  group_by(Claimant) %>%
  summarize(APcheck = n())

TG = checked_claims %>%
  filter(TG == "1") %>%
  group_by(Claimant) %>%
  summarize(TGcheck = n())

checkers = left_join(PF
                     , left_join(WP
                                 , left_join(FC,
                                             left_join(NYT,
                                                       left_join(CNN
                                                                 , left_join(AP, TG, 
                                                                             by = "Claimant"))))))
checkers                                                                       
```

