---
title: "Third Obama-Romney Presidential Debate"
subtitle: "Claimbuster Data, EDA"
output: pdf_document
---

```{r}
library(rvest)
library(dplyr)
library(tidyr)
library(tidyverse)
```


```{r}
all_sentences = read_csv("data/all_sentences.csv")
crowdsourced = read_csv("data/crowdsourced.csv")
groundtruth = read_csv("data/groundtruth.csv")
```

```{r}
rom_ob_all = all_sentences %>% 
  filter(File_id == "2012-10-22.txt")

rom_ob_crowd = crowdsourced %>% 
  filter(File_id == "2012-10-22.txt")

rom_ob_truth = crowdsourced %>% 
  filter(File_id == "2012-10-22.txt")
```

```{r}
rom_ob_entire <- rom_ob_all %>%
  filter(Speaker == "Mitt Romney" |
         Speaker == "Barack Obama")
```


```{r}
full_join(rom_ob_crowd, rom_ob_truth, Sentence_id = Sentence_id)
#essentially, these datasets are identical
write.csv(rom_ob_truth, 'rom_ob.csv', row.names=FALSE) 
```


**README Info from Claimbuster Datasets**

The ClaimBuster dataset consists of six files, the three used are the following:
groundtruth.csv, (22,501 sentences)
crowdsourced.csv, (1032 sentences)
all_sentences.csv, (32,072 sentences)

Both groundtruth.csv and crowdsourced.csv files contain the following attributes.
- Sentence_id: A unique numerical identifier to identify sentences in the dataset.
- Text: A sentence spoken by a debate participant. 
- Speaker: Name of the person who verbalized the Text.
- Speaker_title: Speaker’s job at the time of the debate.
- Speaker_party: Political affiliation of the Speaker.
- File_id: Debate transcript name.
- Length: Number of words in the Text.  
- Line_number: A numerical identifier to to indicate the order of the Text in the debate transcript.
- Sentiment: Sentiment score of the Text. The score ranges from -1 (most negative sentiment) to 1 (most positive sentiment). 
- Verdict: Assigned class label (1 when the sentence is CFS,0 when the sentence is UFS, and -1 when sentence is NFS).

all_sentences.csv file contains all presidential debate sentences. It has all the features shown above except for ``Verdict''. It also includes the following attribute:
- Speaker_role: It depicts the role of the Speaker in the debate as a participant.

**FACT CHECKED DATA SET**
```{r}
ob_rom_factcheck <- read.csv("data/ob_rom_factcheckedset.csv")
```


