---
title: "Third Obama-Romney Presidential Debate"
subtitle: "Claimbuster Data, EDA"
output: pdf_document
---

```{r load-packages, echo=FALSE, message=FALSE, warning=FALSE}
library(rvest)
library(dplyr)
library(tidyr)
library(tidyverse)
library(knitr)
```

```{r load data, message=FALSE}
all_sentences = read_csv("data/all_sentences.csv")
crowdsourced = read_csv("data/crowdsourced.csv")
groundtruth = read_csv("data/groundtruth.csv")
```

```{r load files, echo=FALSE}
rom_ob_all = all_sentences %>% 
  filter(File_id == "2012-10-22.txt")

rom_ob_crowd = crowdsourced %>% 
  filter(File_id == "2012-10-22.txt")

rom_ob_truth = crowdsourced %>% 
  filter(File_id == "2012-10-22.txt")
```

```{r, echo=FALSE}
rom_ob_entire <- rom_ob_all %>%
  filter(Speaker == "Mitt Romney" |
         Speaker == "Barack Obama")
```

```{r, echo=FALSE}
full_join(rom_ob_crowd, rom_ob_truth, Sentence_id = Sentence_id)
#essentially, these datasets are identical
write.csv(rom_ob_truth, 'rom_ob.csv', row.names=FALSE) 
```

**README Info from Claimbuster Datasets**

The ClaimBuster dataset consists of six files, the three used are the following: groundtruth.csv, (22,501 sentences) crowdsourced.csv, (1032 sentences) all_sentences.csv, (32,072 sentences)

Both groundtruth.csv and crowdsourced.csv files contain the following attributes. - Sentence_id: A unique numerical identifier to identify sentences in the dataset. - Text: A sentence spoken by a debate participant. - Speaker: Name of the person who verbalized the Text. - Speaker_title: Speaker's job at the time of the debate. - Speaker_party: Political affiliation of the Speaker. - File_id: Debate transcript name. - Length: Number of words in the Text.\
- Line_number: A numerical identifier to to indicate the order of the Text in the debate transcript. - Sentiment: Sentiment score of the Text. The score ranges from -1 (most negative sentiment) to 1 (most positive sentiment). - Verdict: Assigned class label (1 when the sentence is CFS,0 when the sentence is UFS, and -1 when sentence is NFS).

all_sentences.csv file contains all presidential debate sentences. It has all the features shown above except for \`\`Verdict''. It also includes the following attribute: - Speaker_role: It depicts the role of the Speaker in the debate as a participant.

### **FACT CHECKED DATA SET**

```{r, echo=FALSE}
ob_rom_factcheck <- read.csv("data/obrom_checkeddata.csv")

ob_rom_factcheck <- ob_rom_factcheck %>%
  mutate(Checked = Checked.)
```

```{r, echo=FALSE}
ob_rom_factcheck %>%
  filter(Checked == 1) %>%
  group_by(Speaker) %>%
  summarize(count = n()) %>%
  kable(caption = "# of Checked Claims")
```

Total claims checked: 105

```{r, echo=FALSE}
ob_rom_factcheck %>%
  count(Checked, Speaker) %>%
  pivot_wider(id_cols = c(Checked, Speaker),
              names_from = Speaker, 
              values_from = n, 
              values_fill = 0) %>%
  kable(caption = "Unchecked versus Checked Claims")
```

13% of Obama's statements were checked, compared to 12% of Romney's.

*No apparent partisan bias*

```{r, echo=FALSE}
ob_rom_factcheck <- ob_rom_factcheck %>%
  mutate(Checkability = 
        case_when(Verdict == "-1" ~ "Not Factual Statement",
                  Verdict == "0" ~ "Uncheckworthy Factual Statement",
                  Verdict == "1" ~ "Checkworthy Factual Statement"))

ob_rom_factcheck %>%
  count(Checkability, Speaker) %>%
  pivot_wider(id_cols = c(Checkability, Speaker),
              names_from = Speaker, 
              values_from = n, 
              values_fill = 0) %>%
  kable(caption = "Checkability Score for All Claims")

ob_rom_factcheck %>%
  filter(Checked == 1) %>%
  count(Checkability, Speaker) %>%
  pivot_wider(id_cols = c(Checkability, Speaker),
              names_from = Speaker, 
              values_from = n, 
              values_fill = 0) %>%
  kable(caption = "Checkability Score for Checked Claims")
```

*Of all claims checked:* 52.1% of Obama's claims were Checkworthy Factual Statements, compared to 45.6% of Romney's. 29.2% of Obama's claims were Not Factual Statements, compared to 26.3% of Romney's. 18.8% of Obama's claims were Uncheckworthy Factual Statements, compared to 28.1% of Romney's.

**Of claims checked, the most notable difference was in Uncheckworthy Factual Statements, with about 10% more of Romney's UFS checked.**

*Of all total claims:* 29.5% of Obama's claims were Checkworthy Factual Statements, compared to 23.6% of Romney's. 55.2% of Obama's claims were Not Factual Statements, compared to 67.1% of Romney's. 15.3% of Obama's claims were Uncheckworthy Factual Statements, compared to 9.3% of Romney's.

**The distribution of total claims is much different than that of checked claims. Checkworthy Factual Statments make up a significantly smaller percentage of total claims than checked claims, which makes logical sense. The greatest partisan difference among total claims is that Romney made about 12% more Not Factual Statements than Obama, and Obama made about 1.5x as many Uncheckworthy Factual Statements as Romney's.**

*Comparing checked claims to total claims:* Obama 23.6% of Checkworthy Factual Statements were checked. 7.1% of Not Factual Statements were checked. 16.3% of Uncheckworthy Factual Statements were checked.

Romney 23.2% of Checkworthy Factual Statements were checked. 4.7% of Not Factual Statements were checked. 36.4 % of Uncheckworthy Factual Statements were checked.

**Obama and Romney had an extremely similar percentage of CFS checked. Obama had a higher percentage (by about 1.5x) of his Non Factual Statements checked. Romney had a much higher percentage (about 2x) of his Uncheckworthy Factual Statements Checked.** It may be worth further investigating if this shows a partisan bias.

```{r, echo=FALSE}
ob_rom_factcheck %>%
  group_by(Speaker) %>%
  summarize(avg = mean(Verdict)) %>%
  kable(caption = "More or Less Factual Statements?")
```

With -1 meaning Not Factual Statement, 0 meaning Uncheckworthy Factual Statement, and 1 meaning Checkworthy Factual Statement, I can conclude that in this debate, both candidates made more Nonfactual statements. This data indicates that Obama may make more Factual Statements than Romney.

```{r, echo=FALSE}
ob_rom_factcheck %>%
  filter(times.fact.checked != "#REF!") %>%
  count(times.fact.checked, Speaker) %>%
  pivot_wider(id_cols = c(times.fact.checked, Speaker),
              names_from = Speaker, 
              values_from = n, 
              values_fill = 0) %>%
  kable(caption = "Instances of Repeated Checks")
```

**The distribution of number of times a claim was checked by speaker is relatively equal, Romney had significantly more unchecked statements as he originally made more statements than Obama.**

```{r, echo=FALSE, message=FALSE}
CNN <- ob_rom_factcheck %>%
  filter(CNN == "1") %>%
  group_by(Speaker) %>%
  summarize(CNN_check = n())

Reuters <- ob_rom_factcheck %>%
  filter(Reuters == "1") %>%
  group_by(Speaker) %>%
  summarize(Reuters_check = n())

APNews <- ob_rom_factcheck %>%
  filter(AP.News == "1") %>%
  group_by(Speaker) %>%
  summarize(AP_check = n())

NYT <- ob_rom_factcheck %>%
  filter(NYT == "1") %>%
  group_by(Speaker) %>%
  summarize(NYT_check = n())

Atl <- ob_rom_factcheck %>%
  filter(Atlantic == "1") %>%
  group_by(Speaker) %>%
  summarize(Atl_check = n())

PF <- ob_rom_factcheck %>%
  filter(Politifact == "1") %>%
  group_by(Speaker) %>%
  summarize(PF_check = n())

WP <- ob_rom_factcheck %>%
  filter(WashPost == "1") %>%
  group_by(Speaker) %>%
  summarize(WP_check = n())

FC <- ob_rom_factcheck %>%
  filter(Fact.Check == "1") %>%
  group_by(Speaker) %>%
  summarize(FC_check = n())

checkers = left_join(CNN, 
                     left_join(Reuters, 
                               left_join(APNews, 
                                         left_join(NYT, 
                                                   left_join(Atl, 
                                                             left_join(PF, left_join(WP, FC, by = "Speaker")))))))

checkers %>%
  kable(caption = "Fact Checks by Publisher")
```

1/8 fact checkers checked both speakers equally. (PolitiFact) 3/8 fact checkers checked Obama more than Romney. (Reuters, the Atlantic, the NYT) 4/8 fact checkers checked Romney more than Obama. (CNN, AP News, the Washington Post, Fact Check)
